---
title: "quaker manual"
author: "Ian Towey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **_quaker_** API allows users to retrieve, analyse and visualize global seismic activity. The data used is available on the [Earthquake Hazards Program](http://earthquake.usgs.gov/earthquakes/feed/v1.0/geojson.php) sections of the USGS website.

## quaker package info

The `quaker` API exposes 6 functions.

1. get_seismic_data(timeFrame, minMagnitude)
  * returns a [GeoJson](http://geojson.org/) formatted document of global seismic activity
2. get_freq_grp_by_day_mag(geoJson)
  * Returns summary information ina dataframe of frequency grouped by magnitude by day
3. get_freq_grp_by_country_mag(geoJson)
  * returns a dataframe of frequency grouped by magnitude by country
4. flatten_to_table(geoJson)
  * returns a data frame containing fields 'longtiude','latitude','depth','magnitude','tsunami' , 'time'
5. fit_quaker(geoJson)
  * fits a logistic regression model $$tsunami \sim longtiude + latitude + depth + magnitude$$, returns a list of objects including model fit and anova objects
6. plot(geoJson)
  * Plots an object of type seismic_geojson. Using leaflet creates map of the world with pins at the correct latitude/longtitude for each earthquake in the geoJson object. Each pin has a toop tip which when clicked provides futher information on the seismic event.



## Example 1

Download all the seismic events records by the USGS for the past month with magnitude greater than 2.5 on the Richter scale.
Group the the frequency by country and magnitude and order by total number of seismic events per country descending.
```{r, echo=FALSE, include=FALSE}
library(quaker)
library(dplyr)
```

Gr
```{r}
data <- get_seismic_data(timeFrame = 'PAST_MONTH', minMagnitude = '2.5')
grp_by_cntry_mag <- get_freq_grp_by_country_mag(data)
grp_by_cntry_mag = grp_by_cntry_mag %>% arrange(desc(total)) %>% head
```

```{r, echo=FALSE, results='asis'}
knitr::kable(grp_by_cntry_mag)
```

## Example 2

Download all the seismic events records by the USGS for the past week with any magnitude. Group the the frequency by day and
magnitude and order by total number of seismic events per day descending.


```{r}
data <- get_seismic_data(timeFrame = 'PAST_WEEK', minMagnitude = 'all')
grp_by_day_mag <- get_freq_grp_by_day_mag(data)
```

```{r, echo=FALSE, results='asis'}
knitr::kable(grp_by_day_mag)
```

## Example 3

Download all the seismic events records by the USGS for the past week with any magnitude. Flatten the Geojson object to data frame for ease of manupulation
Returning the fields

  + _longtiude_ : Decimal degrees longitude. Negative values for western longitudes.
  + _latitude_  : Decimal degrees latitude. Negative values for southern latitudes.
  + _depth_     : Depth of the event in kilometers.
  + _magnitude_ : The magnitude for the event.
  + _tsunami_   : Indicator warning indicator
  + _time_      : Time when the event occurred. Times are reported in milliseconds since the epoch

```{r}
data <- get_seismic_data(timeFrame = 'PAST_WEEK', minMagnitude = 'all')
data_flat <- flatten_to_table(data)
```

```{r, echo=FALSE, results='asis'}
knitr::kable(head(data_flat))
```

## Example 4

Download all the seismic events records by the USGS for the past month with any magnitude. Fit a logistic regression
$$tsunami \sim longtiude + latitude + depth + magnitude$$

  ```{r}
data <- get_seismic_data(timeFrame = 'PAST_MONTH', minMagnitude = 'all')
model.fit <- fit_quaker(data)

```

```{r}
model.fit$fit
```

```{r}
model.fit$fit.anova
```

## Example 5

Download all the seismic events records by the USGS for the past week with magnitude greater than 2.5. Plot map of world with pins showsing event location
and pin tooltip with more details of the event


```{r, eval=FALSE}
this_week_2_5 <- get_seismic_data(timeFrame = 'PAST_WEEK', minMagnitude = '2.5')
plot(this_week_2_5)
```
